Distribution of a Facebook application

### 2011-04-10 paper.tex
first version completed

When clients access the canvas page of the application, they are
redirected to the OAuth Dialog (running on a Facebook server) and
provided with the application's ID. Then, user authentication and
app authorization is performed, by involving the client and Facebook
only. User authentication is done by logging in on Facebook (if necessary) and
authorization of the application is done by asking the user if he/she
wants to grant the application access to some data specified by the
application with the scope parameter. One should note, that no
login data are sent to the application, thus, this is how Facebook
establishes a trust relation. 
If the user authorizes the application, an \textit{authorization code} is returned and
the user is redirected to the application, to which the
\textit{authorization code} is forwarded. When registering the
application, an \textit{application secret} code has been
created. In order to authenticate the application, the
\textit{authorization code} and the \textit{application secret} have
to be sent to Facebook's \textit{Graph API Token Endpoint}. Facebook
then returns an \textit{access token}, which can be included in the
URL by which the Graph API can be accessed. However, Facebook's
\ac{sdks} also provide functions to perform calls to the Graph API,
and in this case, the access token is included automatically.

@@

\begin{figure}
  \includegraphics[width=3.5in]{../img/fbsolution.png}
  \caption{Solution of a question.}
\label{fig:solution}
\end{figure}

@@

\subsection{Data evaluation}
\label{sec:data-evaluation}
Data evaluation is also performed by using the MySQL tables, and
PHP. In addition, the dot
language\footnote{http://www.graphviz.org/Documentation.php} is used
to create graphs.

The evaluation scripts are maintained by \textit{main.php}. A user
class represents Facebook's social graph internally. Properties of
user objects equal the \textit{user}-table's properties, and a user's
friendships are represented by a list of references to corresponding
user objects. The whole set of users is maintained by
\textit{main.php}, where they are stored in an array. Class \textit{Graph}
allows for output of various graphs in dot-language created from the
internal social graph.

\section{Results}
\label{sec:results}

@@

\section{Discussion}
\label{sec:discussion}

\section{Outlook}
\label{sec:outlook}

%\section{Conclusion}
%\label{sec:conclusion}

@@

>>>>

When clients access the canvas page of the application, they are
redirected to the OAuth Dialog (running on a Facebook server) and
provided with the application's ID. Then, user authentication and
app authorization is performed, by involving the client and Facebook
only. User authentication is done by logging in on Facebook (if necessary) and
authorization of the application is done by asking the user if he/she
wants to grant the application access to some data specified by the
application with the scope parameter. One should note, that no
login data are sent to the application, thus, this is how Facebook
establishes a trust relation. 
If the user authorizes the application, an \textit{authorization code} is returned and
the user is redirected to the application, to which the
\textit{authorization code} is forwarded. During registration of the
application, an \textit{application secret} code was
created. In order to authenticate the application, the
\textit{authorization code} and the \textit{application secret} have
to be sent to Facebook's \textit{Graph API Token Endpoint}. Facebook
then returns an \textit{access token}, which can be included in the
URL by which the Graph API can be accessed. However, Facebook's
\ac{sdks} also provide functions to perform calls to the Graph API,
and in this case, the access token is included automatically.

\subsection{Graphs with DOT language}
\label{sec:graph-repr-with}
The \ac{dot} language allows for plain text description of
graphs. There are few programs that can process \ac{dot} graphs, most
of them are part of the
\textit{Graphviz}\footnote{http://www.graphviz.org}-package. 
The language allows for directed and undirected graphs, attributes can
be set and colors as well as different shapes for nodes and edges can
be used. The basic syntax is very easy, the graph shown in figure
\ref{fig:dot} has the \ac{dot}-code given in listing \ref{list:dot}.

\begin{lstlisting}[caption=A very simple DOT-Graph,
  label=list:dot]{}{
 digraph myGraph {
     a -> b;
}
\end{lstlisting}

\begin{figure}
  \includegraphics[width=0.4in]{../img/ab.png}
\caption{A very simple graph in \ac{dot}-language.}
\label{fig:dot}
\end{figure}

Labels and colors can be added by defining the node in the head of the
graph description. An example is detailed in listing \ref{list:dot2}, and
the corresponding graph is given in figure \ref{fig:dot2}.

\begin{figure}
  \includegraphics[width=0.4in]{../img/graphDot.png}
\caption{A very simple graph with node a blue boxed and fontcolor set
  to red.}
\label{fig:dot2}
\end{figure}

\begin{lstlisting}[caption=A very simple DOT-Graph,
  label=list:dot2]{}{
 digraph myGraph {
a [label="Foo", color="blue", shape=box, fontcolor="red"];
     a -> b;
}
\end{lstlisting}

@@

@@

\subsection{Data evaluation}
\label{sec:data-evaluation}
Data evaluation is performed by using the MySQL tables, and
PHP. In addition, the dot
language\footnote{http://www.graphviz.org/Documentation.php} is used
to create graphs.

The evaluation scripts are maintained by \textit{main.php}. A user
class represents Facebook's social graph internally. Its properties
are user's Facebook ID, registration date, average score, a list of
friends, and a list of IDs to whom the user may have spread the
application. Friendships are represented by a list of references to corresponding
user objects, and names of users' can be made anonymous if needed.

A user is supposed to have spread the app if he/she has
a friend who used the application and who belongs to a \textit{new}
network. Such persons are found by recursively checking if an user of
the application has friends, who also used the application, and whose friends
also used the app, but are not in his/her list of friends, and also not in the list of the person
who originally spread the application. In this case, the corresponding
friend will be marked as a person who spread the application, and the
recursion starts by letting the friend become the analyzed user.
Since it is not clear, why a person started to use an application, the
script also allows for identification of users who have friends 
who also participated in the application.

User names are made anonymous by substituting their names with a
random name selected from a list of popular baby names
from 2009, taken from \url{http://www.ssa.gov/cgi-bin/popularnames.cgi}.   
For each gender, hundred names are stored in an array, and if the
gender is set, a name of same gender is chosen. Otherwise, the name is
taken randomly. The name of the person who spread the application
originally is always set to \textit{Alexandra}.

Output options are a \ac{csv}-file containing metrics calculated from
the internal social graph as well as various graphs in dot language, created
from the internal social graph, too.
% Properties of
% user objects are in the style of the \textit{user}-table's properties, and a user's
% . The whole set of users is maintained by
% \textit{main.php}, where they are stored in an array. Class \textit{Graph}
% allows for output of various graphs in dot-language created from the
% internal social graph.

\section{Results}
\label{sec:results}
A total of n=48 users participated in the quiz. 5505 users have been
stored in the user table. Thus, on average each user brought on
average 114.7 \textit{new} people, who could read about the application
in their news stream. 15

\subsection{Characteristics}
\label{sec:characteristics}

The characteristics discussed in this section consider data of users
that are still registered (n=37).

\subsubsection{Age and gender}
\label{sec:age-gender}
$59 \%$ of the registered user are female, $41 \%$ of the participants
are male. Table \ref{tab:agegender} shows classification of the users
by age and gender. It can be seen, that most users ( 70 \%) were at
the age of 25-34. Only few ($8.4 \%$) were older, and only 21.4 \%
were younger.

\begin{center}
\begin{table}
\label{tab:agegender}
\begin{tabular}{l | c c c c}
Gender\textbackslash Age  & 13-17 & 18-24 & 25-34 & 45-54 \\
\hline
female  & 2.7 \% & 16 \% & 35 \% & 5.4 \% \\
male & 0 \% & 2.7 \% & 35 \% & 2.7 \% 
\end{tabular}
\caption{Table shows classification of users by age and gender.}
\end{table}
\end{center}

\subsubsection{Demography}
\label{sec:demographie}
Table \ref{tab:countries} details the origin of the users and table
\ref{tab:languages} the languages in which the participants use Facebook.
As it seems obvious for a quiz in German, most visitors came from Germany.

\begin{center}
\begin{table}
\label{tab:countries}
\begin{tabular}{l | c }
Country  & Visitors \\
\hline
Germany  & 32 \\
United Kingdom & 2 \\
Israel & 1 \\
United States of America & 1 \\
Austria & 1 \\
\end{tabular}
\caption{Table shows where the users are from.}
\end{table}
\end{center}

\begin{center}
\begin{table}
\label{tab:languages}
\begin{tabular}{l | c }
Languages   & Visitors \\
\hline
German  & 25 \\
English (UK) & 7 \\
English (US) & 4 \\
Greek  & 1 \\
\end{tabular}
\caption{Languages of the participants.}
\end{table}
\end{center}

\subsection{Spreading of the quiz}
\label{sec:distribution-quiz}
In this section, graphs illustrating how the
application spread are presented. The original user names collected from Facebook were made anonymous
before the graphs were printed (see section \ref{sec:data-evaluation}). Friendships are
represented by directed edges: If user A has user B in the friend list, an
arrow exists targeting user B.

Figure \ref{fig:allUsers} represents each user stored in user
table as node, and friendships of those who participated in the
application as directed edges. Names of users who participated in the quiz are distinguished by
red font color. The main purpose of this figure is to visualize the
whole setting and to give an imagination of how many \textit{new}
people on Facebook could have read about the application in their 
news stream. In addition, the figure shows few participants, who are
not connected to the larger network of the 
person who started to spread the application.

\begin{figure}
  \includegraphics[width=5.2in]{../img/allUserTree.png}
\caption{Figure shows all users stored in the database. Users who
  participated have a red text color.}
\label{fig:allUsers}
\end{figure}

In the following, only user who participated in the quiz are
considered. In Figure \ref{fig:spreading} nodes of people, who are
connector between two networks, have a red font color. The number of
friends and the average score from quiz are also shown. 
It can be obtained that in three cases, there are several connections (friendships)
between the people who participated and belong to a new network. For
instance, several connections exist between users \textit{Evan},
\textit{Diego}, \textit{Isabelle}, \textit{Chase}, and \textit{Gebrielle}.
\textit{Nicholas} and \textit{Chase}, who are friends with those
people mentioned above, also are friends with \textit{Ethan},
and \textit{Josiah}.  Then, \textit{Faith} is friend with
\textit{Owen} and \textit{Colton}, who also know each other, and
\textit{Mariah} and \textit{Sydney} have \textit{Alexis} as a common
friend. The figure also shows that most people who spread the
application had a score greater than or equal to three. At least, it
can be obtained that half of the people who spread the application
had less than hundred friends, which is less than the mean on
Facebook, i.e. 130 friends \cite{ref:facebookStats}.

In the addendum (section \ref{sec:addendum}), two further graphs are
shown. In figure \ref{fig:friendsScored}, the background color
represents a user's average score. User with a darker background had a
higher score with those having a lighter background color. 
Figure \ref{fig:friendsFilled} visualizes the amount of friends in the
by giving users having many friends a dark and those having fewer
friends a lighter background color.

\begin{figure}
  \includegraphics[width=5in]{../img/spreading.png}
\caption{Figure shows all users stored in the database. Users who
  participated have a red text color.}
\label{fig:spreading}
\end{figure}

Two tables also located in the addendum describe to how many
other participants friendships existed (see table \ref{tab:spread} and
\ref{tab:nospread}), and to how many people of a
\textit{new} network the application was (probably) distributed by the
user. $26$ participants were on average $2.54$ friends with other
participants and those, who spread the application, seemed to have
more connections to other people than those who did not.
The average score of people connected to others was $3.02$. It was
higher than the average score of people not connected to others ($2.41$).
People not connected to further participants had fewer friends on average ($106.8$)
than the connected participants ($138.83$).

Table \ref{tab:nouser} details how many friends who participated in
the application user had, who did not participate in the
application. Most times, those people just had one friend, but in eleven
cases, they had four or five participants in their friends list.

\begin{table}
\label{tab:nouser}
\begin{tabular}{c c}
Number of friends & Frequency \\
1& 5150\\
2& 240\\
3& 55\\
4& 6\\
5& 5\\
\end{tabular}
\caption{Table details how many friends user had, who did not
  participate in the application.}
\end{table}

@@

\section{Discussion and Outlook}
\label{sec:discussion}
From the results given in section \ref{sec:results} it can be derived
that the number of friends did not to play a bigger role in the
process of the application's distribution. The number of connections
to other users seems to play a bigger 
role, as it is indicated by table \ref{tab:spread} and figure
\ref{fig:spreaders}. In addition, a worse score made people unlikely
to spread the application, just one person with a score less than 3
did so. Thus, people with a higher score (on average $3.11$, median $3.32$) are
supposed to have distributed the application. Indeed, some people who
had four or five friends who participated in the application did not
install the application.

In order to get more reliability towards the processes of distribution
of applications on Facebook, more applications must be analyzed.
Future research may consider, how patterns change with respect
to users' age, and if further characteristics can be identified. In
case of larger applications, it could be studied, if turning points
or intervals exist, which - when exceeded - make users significantly
more likely to join the application.

\section{Conclusion}
\label{sec:conclusion}
This research indicates, that the spreading of application follows
certain patterns. Oftentimes, people say that they don't want to
make a worse impression on Facebook, and this is also reflected in the
fact, that only people who yield higher scores in the quiz also
distributed to their friends. In the environment analyzed, dense
networks with few participants, i.e. in networks in which people know
each other, made  further members of their network very likely to join
the application. Indeed, the number of friends did not play a role in
making people likely to spread the application.

@@

\newpage
\section{Addendum}
\label{sec:addendum}
\begin{figure}
  \includegraphics[width=3.5in]{../img/fbsolution.png}
  \caption{Solution of a question.}
\label{fig:solution}
\end{figure}

\begin{figure}
  \includegraphics[width=5in]{../img/friendsScored.png}
\caption{Figure shows all users stored in the database. Users who
  participated have a red text color.}
\label{fig:friendsScored}
\end{figure}

\begin{figure}
  \includegraphics[width=5in]{../img/friendsFilled.png}
\caption{Figure shows all users stored in the database. Users who
  participated have a red text color.}
\label{fig:friendsFilled}
\end{figure}

 \begin{table}
 \label{tab:spread}
 \begin{tabular}{ c c c c}
Friends & Score & Friends who also participated & Probably spread appl. to friends\\
\hline
97 & 3.33 & 1 &-\\
322 & 1.5 & 4 & -\\
63 & 4 & 1 & 1\\
102 & 0 & 2 & -\\
91 & 1 & 3 &1\\
128 & 4 & 2 & -\\
151 & 3 & 4 &3\\
135 & 3.64 & 3 &2\\
254 & 4 & 2 & 2\\
220 & 3 & 4 & -\\
127 & 3 & 2 & -\\
186 & 4 & 3 & -\\
42 & 3 & 4 & - \\
445 & 3 & 4 & - \\
153 & 3 & 2 &1\\
20 & 3 & 1 &1\\
11 & 4.23 & 8 &2\\
130 & 3 & 1 & -\\
67 & 2 & 1 & - \\
207 & 2 & 2 & -\\
74 & 4.5 & 3 & -\\
47 & 4.42 & 3& -\\
157 & 3 & 2 & -\\
136 & 3 & 2 & -\\
78 & 4 & 1 & -\\
167 & 2 &1 & -
\end{tabular}
\caption{Table shows how many friends and which score user had who may have spread the
  application, and how many of their friends started to use the
  application. If a user was a connector to another network, it is
  calculated to how many new persons the application (appl.) was
  probably spread.}
\end{table}

 \begin{center}
 \begin{table}
 \label{tab:nospread}
 \begin{tabular}{ c c c}
Friends & Score  \\
\hline
90 & 1.29 \\
332 & 3 \\
233 & 4 \\
79 & 3 \\
53 & 1 \\
89 & 2 \\
72 & 4 \\
76 & 1 \\
264 & 1 \\
35 & 3 \\
121 & 4 \\
289 & 1 \\
15 & 4 \\
49 & 4 \\
0 & 4.75 \\
80 & 4 \\
60 & 0 \\
13 & 2.5 \\
229 & 0 \\
53 & 0 \\
10 & 3 
\end{tabular}
\caption{Table shows how many friends and which score user had who did
  not spread the application.}
\end{table}
\end{center}
